---
title: "Class 13: Structure Based Drug Discovery"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Download and clean up the receptor structure

Here we will download the 1HSG PDB structure from the online database, remove the bound drug Merck and write out a protein only and ligand only new set of PDB files

We will use the bio3d package function `get.pdb()`, `read.pdb()`, `atom.select()`, and `write.pdb()`

```{r}
# Download the PDB structure

library(bio3d)

file.name <- get.pdb("1HSG")
hiv <- read.pdb(file.name)
hiv
```

Q1: What is the name of the two non protein resid values in this structure? What does resid correspond to and how would you get a listing of all reside values in this structure? 

Water (HOH) and Merck (MK1, ligand); resid corresponds to the residue name within the protein

```{r}
prot<-atom.select(hiv, string="protein", value=TRUE)
write.pdb(prot, file="1hsg_protein.pdb")
```

Select and write out ligand file only

```{r}
lig <- atom.select(hiv, string="ligand", value=TRUE)
write.pdb(lig, file="1hsg_protein.pdb")
lig
```

