---
title: "Class 12: Structural Bioinformatics 1"
output: github_document
---

# PDB Statistics

Here we inspect the types of structures in the main stabase for 3D biomolecular data - the PDB.

Q1: Determine the percentage of structures solved by X-Ray and Electron Microscopy. Also can you determine what proportion of structures are protein?

```{r}
# read in the file
stats <- read.csv("Data Export Summary.csv", row.names = 1)
stats
# calculate the percentages asked for - practice
(stats$Proteins[1]) / (sum(stats$Total))

(stats$Proteins[3]) / (sum(stats$Total))

# calculate the percentage values for each method - answer
sum(stats$Total)

ans<-(stats$Total)/(sum(stats$Total))*100
names(ans)<-row.names(stats)
round(ans,2)

# calculate the percentage values for proteins - answer
round(sum(stats$Proteins)/sum(stats$Total)*100,2)
```

Q2: Type HIV in the PDB website search box on the home page and determine how many HIV-1 protease structures are in the current PDB?
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Including Code

You can include R code in the document as follows:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
