---
title: "Class7"
author: "Sarah Murphy"
date: "1/28/2020"
output: ioslides_presentation
---
#WRONG OUTPUT - SEE CLASS-7.RMD
##Class 7
###Recap: How to Use Functions
Q6 Homework: Make sure you understand what each line of code does. Must be a separate file from in class activities in an R Markdown file.

Structure: `fname <- function(arg1,arg2) {paste(arg1, arg2)}`
Input arguments: predefine arguments - call it 
Function body: arguments do whatever you tell them to do in the function body

Example: `rescale <- function(x, na.rm=TRUE, plot=FALSE, ...) {
 rng <-range(x, na.rm=na.rm)
 answer <- (x - rng[1]) / (rng[2] - rng[1])
 if(plot) {
 plot(answer, ...)
 }
 return(answer)
}'

rescale = function

For multiple conditions, use "if" statements: branching logic trees to return different values for different conditions.

##Process for Coding a Function
1. Start with a simple problem and write a working snippet of code.
2.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

