---
title: "Class 6 - R Function"
author: "Sarah Murphy"
date: "1/23/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Functions

##Level 2 heading

###Level 3 Heading

#Notes
#Data Import
Read.table(“your file name”, header = TRUE, Sep = “\\t”)

```{r}
#This is a silly plot
plot(1:5)
```

Let's see more about **file import** (i.e. reading files into R) The main read function in base R is `read.table()`
Use different read functions for different types of data. They each have different predefined arguments that are specific for different types of data. This saves typing in creating code.

```{r}
read.table("test1.txt")
```

We need to add arguments to get this file read properly!

```{r}
read.table("test1.txt",sep = ",", header = TRUE)
```

This is how to read this type of data properly. But this could also be done using `read.csv`, which has the arguments I want in this case!

```{r}
t1<-read.csv("test1.txt")
t1
```

The second file
```{r}
t2<-read.table("test2.txt", header = TRUE, sep = "$")
t2
```

The third file
```{r}
t3<-read.table("test3.txt")
t3
```

# Back to functions

Our first example function

```{r}
add <- function(x, y=1) {
 # Sum the input x and y
 x + y
}
```

This function has now been added to our global environment. Let's try using it!

```{r}
add(34)
```
```{r}
add(32,35)
```

How does this work with vectors as the input?

```{r}
add( c(1, 2, 3) )
add( c(1, 2, 3), 4 )
```

How about multiple vectors?

```{r}
add( c(1, 2, 3), c(1,2,3) )
```

If it is not made into a vector, the additional numbers are considered extra arguments, and the function does not know what to do with them. This will result in an error message.
```{r}
#add( c(1, 2, 3), 1,2,3 )
```

# Writing Your Own Functions

Helps simplify work and clarify purpose of code
Start with a working code snippet, simplify your variables, reduce calculation duplication, and *finally* turn it into a function!
i.e. df$a = x
`minx <- min (x)`
`rng<-range(x)`

What is this `range()` function you talk of?

```{r}
x<-c(4,4,10,3,11)
max(x)
min(x)
range(x)
```

This is our second function
```{r}
rescale <- function(x) {
 rng <-range(x)
 (x - rng[1]) / (rng[2] - rng[1])
}
rescale(x)
```

Let's do some more testing of our rescale function:
```{r}
rescale(1:10)
```

What happens when a data value has NA in it?
```{r}
rescale( c(1,2,NA,3,10) )
```
The output is all "NA"! This is not what we want.

```{r}
x<-( c(1,2,NA,3,10) )
rng <-range(x)
rng
```

```{r}
rng <-range(x, na.rm=TRUE)
rng
```

```{r}
rescale2 <- function(x) {
 rng <-range(x, na.rm = TRUE)
 (x - rng[1]) / (rng[2] - rng[1])
}
rescale2( c(1,2,NA,3,10) )
```

Going too far

```{r}
rescale3 <- function(x, na.rm=TRUE, plot=FALSE) {
 rng <-range(x, na.rm=na.rm)
 print("Hello")
 answer <- (x - rng[1]) / (rng[2] - rng[1])
 print("is it me you are looking for?")
 if(plot) {
   print("Don't sing again!")
 plot(answer, typ="b", lwd=4)
 }
 print("I can see it in ...")
 return(answer)
}
```

```{r}
rescale3(x, plot = FALSE)
```

#Hands On Worksheet
##Section B

```{r}
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b

plotb3(s1.b, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```

#Q1. What type of object is returned from the read.pdb() function?
`read.pdb()` reads a Protein Data Bank coordinate file and pulls up the data into the Global Environment. By doing this, the data can be named and easily input into functions to analyze. `read.pdb()` returns a list of 8 things and of class "pdb", "sse".
#Q2. What does the trim.pdb() function do?
`trim.pdb()` trims a PDB object to a subset of atoms. 
#Q3. What input parameter would turn off the marginal black and grey rectangles in the plots and what do they represent in this case?
The sse parameter turns of the rectangles in the plots. ?plotb3 > dssp tells us that there are 3 values for the rectangles: helix, sheet, or turn.
#Q4. What would be a better plot to compare across the different proteins?
It would be better to have all 3 proteins in one plot so that it is easier to recognize patterns without shuffling between 3 different graphs. This can be done by using `points()` and adding the name
#Q5. Which proteins are more similar to each other in their B-factor trends. How could you quantify this? HINT: try the rbind(), dist() and hclust() functions together with a resulting dendrogram plot. Look up the documentation to see what each of these functions does.
s2.b and s3.b are most closely related. `rbind()` combines vectors by row and `dist()` finds the distance between each protein.

```{r}
s1 <- read.pdb("4AKE") # kinase with drug
s1
```

```{r}
class(s1)
str(s1)
```

```{r}
aa321(s1$seqres)
```

```{r}
s1$atom
```

```{r}
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
points(s2.b, typ="l", col="PURPLE",lwd=2)
points(s3.b, typ="l", col="PINK",lwd=2)

```


```{r}
hc <- hclust( dist( rbind(s1.b, s2.b, s3.b) ) )
plot(hc)
```

#Q6. How would you generalize the original code above to work with any set of input protein structures?
```{r}
library(bio3d)
proteins <- c(read.pdb("4AKE"), read.pdb("1AKE"), read.pdb("1E4Y"))
proteins
```

```{r}
proteins2<-function(proteins, na.rm=TRUE, plot=TRUE) {
 trim <-trim.pdb(proteins, chain="A", elety="CA")
 proteins3<-trim$atom$b
 plot(proteins3, typ="l",col="rainbow",lwd=2)
}
```

