---
title: "Class 10"
author: "Sarah Murphy"
date: "2/6/2020"
output: github_document
---
## PCA: Principal Component Analysis

PCA projects the features onto the principal componenets

The motivation is to reduce the features dimensionality while only losing a small amount of information

The data have a maximum variance along PC1, then PC2, etc., which makes the first few PCs useful for visualizing our data and as a basis for further analysis
• Reduces dimensionality
To visualize multidimensional data
• To choose the most useful variables (features)
• To identify groupings of objects (e.g. genes/samples)
• To identify outliers

### Practical PCA Issue: Scaling

Number of cylinders in engine, mpg, hp, etc. of cars are measured in different scales.
-Must scale data first to treat it equally & compare
-Argument exists in `prcomp()`: scale=FALSE is default --> can set it as true so that data is scaled and centered around the same
  `prcomp(x, scale=TRUE)`
  
# Hands On Worksheet: Unsupervised Mini Project - Analysis of Human Breast Cancer Cells

## First we need to import (i.e. read) our input data

```{r}
wisc.df<-read.csv("WisconsinCancer.csv")
head(wisc.df)
```

There are funky things in this data sat that we will ignore for analysis. Do not include ID, diagnosis or X variables in data analysis --> must know what you're analyzing!


```{r}
# Convert the features of the data: wisc.data

wisc.data<-as.matrix(wisc.df[,3:32])
head(wisc.data)
```

```{r}
# Q1: Number of patients in this dataset

nrow(wisc.data)
```

```{r}
# Create a diagnosis vector for later

diagnosis <- table(wisc.df$diagnosis)
diagnosis

# Q2: There are 212 malignant diagnoses
```

```{r}
# Q3. How many variables/features in the data are suffixed with _mean? --> Use `grep()`
colnames(wisc.data) # By eyeballing, we see there are 10, but we want to use code to ensure we are correct
grep("_mean", colnames(wisc.data))
# By setting value = TRUE in grep function, we can see what the matches to the pattern are
```

```{r}
length(grep("_mean", colnames(wisc.data)))
```

## Principal Component Analysis

Before we turn to PCA, we need to consider whether we shold SCALE our data.

It is important to check if the data need to be scaled before performing PCA. Recall two common reasons for scaling data include:

The input variables use different units of measurement.

- The input variables have significantly different variances.
- Check the mean and standard deviation of the features (i.e. columns) of the wisc.data to determine if the data should be scaled. Use the `colMeans()` and `apply()` functions like you’ve done before.

```{r}
round( apply(wisc.data, 2, sd), 2)
```

Looks like we need to scale = TRUE! The values are different and are clearly measured differently. i.e., some standard deviations are in the 100s, some are less than 1.

```{r}
# Perform PCA on wisc.data by completing the following code
wisc.pr <- prcomp(wisc.data, scale=TRUE)
summary(wisc.pr)

# Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)? 44.27%
# Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data? 3
# Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data? 7
```

Let's make some figures...

```{r}
biplot(wisc.pr)
# This is useless and hard to read.
```

We need to create our own PC1 vs. PC2 plot. Let's color by the diagnosis.
```{r}
attributes(wisc.pr)
```

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=wisc.df$diagnosis, xlab = "PC1", ylab = "PC2")
abline(h=0, col="grey", lty=2)
abline(v=0, col="grey", lty=2)
```

The red (malignant) is spread much farther than the black. It is not a clear relationship, which demands we pay attention and analyze it further.

PCA takes a dataset with a lot of dimensions (i.e. lots of cells) and flattens it to 2 or 3 dimensions so we can look at it.
– It tries to find a meaningful way to flatten the data by focusing on the things that are different between cells. (much, much more on this later)


## Cluster in PC Space

First, let's see if we can cluster the original data

```{r}
wisc.hc<-hclust(dist(wisc.data))
plot(wisc.hc)
```

This does not tell us anything. Rather than input the raw data, let's try combining the results of our PCA with clustering...

Let’s see if PCA improves or degrades the performance of hierarchical clustering.

Using the minimum number of principal components required to describe at least 90% of the variability in the data, create a hierarchical clustering model with the linkage method="ward.D2". We use Ward’s criterion here because it is based on multidimensional variance like principal components analysis. Assign the results to wisc.pr.hclust.

```{r}
# Let's cluster the first 3 PCs
wisc.pr.hclust<- hclust(dist(wisc.pr$x[,1:3]), method="ward.D2")
plot(wisc.pr.hclust)
```

To get our clusters out of this group, we need to CUT it with the `cutree()` function

```{r}
grps3<-cutree(wisc.pr.hclust, k=2)
grps3
table(grps3)
```

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=grps3, xlab = "PC1", ylab = "PC2")
```

We can use the `table()` function to compare the $diagnosis vector with our cluster results vector

```{r}
table(grps3,wisc.df$diagnosis)
```

Overlapping of clusters leads to false positives and negatives --> misclassification of malignant vs. benign tumors.

## Prediction
We will use the predict() function that will take our PCA model from before and new cancer cell data and project that data onto our PCA space.

```{r}
new<- read.csv("new_samples.csv")
new
```

Use th `predict()` function with our previous PCA model and new data...
```{r}
npc<-predict(wisc.pr, newdata=new)
npc
```

Now draw the PCA plot again and add our new data

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=wisc.df$diagnosis)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], c(1,2), col="white")
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
